---
title: "FEMA_Data_Cleaning"
author: "Sean"
date: '2022-07-9'
output:
  pdf: 

---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
### Set Environment ------------------------------------------------------------
# Clear the environment
rm(list=ls())
gc()
# So the code will compile warnings as per usual
options(warn = 0)
# Turn off scientific notation
options(scipen = 999)
### Load Packages --------------------------------------------------------------
if(!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, magrittr, stringr, reshape2, janitor,
lubridate, readxl, data.table, ggplot2, scales, readr,
tidyr, zoo, skimr, openxlsx,ggspatial, rgeos, data.table,RColorBrewer,
tidyverse, rio, collapse, sf, glue, XML, tm, here, purrr, repurrrsive,
tmap,tidygraph, nabor,igraph, viridis, hrbrthemes,RSocrata,soql,xts,
tidycensus,tmaptools, eks,forcats)
```


### Goal of this script is to read the FEMA data and clean it to get the relevant grants during Sandy and Ida.
```{r}
FEMA <- read_csv("../2_Data/FEMA/IndividualsAndHouseholdsProgramValidRegistrations.csv")
names(FEMA)
zips<- read_sf("../2_Data/ZIP_CODE_040114/ZIP_CODE_040114.shp") 

zips_dissolve <- zips %>% 
    group_by(ZIPCODE) %>% 
    summarise(pop = mean(POPULATION)) %>% 
    st_cast() 
    
```
Look at the types of incidents there are:
```{r}
FEMA_incident_Type <- FEMA %>% 
    group_by(incidentType) %>% 
    count() %>% 
    arrange(desc(n)) %>% 
    ungroup() 

FEMA_incident_Type %>% 
    mutate(incidentType = fct_reorder(incidentType, n)) %>% 
    ggplot(aes(y=n, x = incidentType)) +
    geom_bar(stat="identity", fill="#f68060", alpha=.6, width=.4) +
    coord_flip() +
    theme_minimal()

```


Filter down the FEMA dataset so that we're only looking at those that exist in New York City.
```{r}
NYC_FEMA <- FEMA %>% 
    filter(damagedZipCode %in% zips$ZIPCODE)
```

Look at the types of incidents there are:
```{r}
NYC_FEMA_incident_Type <- NYC_FEMA %>% 
    group_by(incidentType) %>% 
    count() %>% 
    arrange(desc(n)) %>% 
    ungroup() 

NYC_FEMA_incident_Type %>% 
    mutate(incidentType = fct_reorder(incidentType, n)) %>% 
    ggplot(aes(y=n, x = incidentType)) +
    geom_bar(stat="identity", fill="#f68060", alpha=.6, width=.4) +
    coord_flip() +
    theme_minimal()
```

What is the time frame for these complaints?
```{r}
NYC_FEMA_overtime <- NYC_FEMA %>% 
    group_by(declarationDate,incidentType) %>% 
    count() %>% 
    ungroup() 

NYC_FEMA_overtime

NYC_FEMA_overtime %>% 
    ggplot(aes(x=declarationDate, y=n, group=incidentType, color=incidentType)) +
    geom_point()

```
Because it looks like there is a very clear distinction of when Sandy and Ida complaints are, we will filter for that date.
They correspond with these dates:

2012-10-30 04:34:00
2021-09-05 23:36:00

```{r}
sandy_date <- as_date("2012-10-30 04:34:00")
ida_date <- as_date("2021-09-05 23:36:00")
hurricane_dates <- c(sandy_date,ida_date)

NYC_FEMA_Hurricane <- NYC_FEMA %>% 
    mutate(date = as_date(declarationDate),
           ihpEligible = ifelse(ihpEligible == 1,"Eligible","Ineligible"),
           floodDamage = ifelse(floodDamage == 1,"damaged","notdamaged")) %>% 
    filter(date %in% hurricane_dates) 
```
```{r}
NYC_FEMA_Hurricane_by_zipcode <- NYC_FEMA_Hurricane %>% 
    group_by(date, damagedZipCode, ihpEligible) %>% 
    count() %>% 
    ungroup() %>% 
    pivot_wider(names_from = c("ihpEligible","date"),values_from = n)

NYC_FEMA_Hurricane_damage_by_zipcode <- NYC_FEMA_Hurricane %>% 
    group_by(date, damagedZipCode, floodDamage) %>% 
    count() %>% 
    ungroup() %>% 
    pivot_wider(names_from = c("floodDamage","date"),values_from = n) %>% 
    rename()

names(NYC_FEMA_Hurricane_by_zipcode)

NYC_FEMA_Spatial <- zips_dissolve %>%
    left_join(NYC_FEMA_Hurricane_by_zipcode, by = c("ZIPCODE"="damagedZipCode")) %>% 
    mutate(eligible_Sandy_pc = `Eligible_2012-10-30`/pop,
           ineligible_Sandy_pc = `Ineligible_2012-10-30`/pop,
           eligible_Ida_pc = `Eligible_2021-09-05`/pop,
           ineligible_Ida_pc = `Ineligible_2021-09-05`/pop,)

NYC_FEMA_damage_Spatial <- zips_dissolve %>%
    left_join(NYC_FEMA_Hurricane_damage_by_zipcode, by = c("ZIPCODE"="damagedZipCode")) %>% 
    mutate(damaged_sandy_pc = `damaged_2012-10-30`/pop,
           damaged_ida_pc = `damaged_2021-09-05`/pop) %>% 
    rename(damaged_sandy = `damaged_2012-10-30`,
           damaged_ida = `damaged_2021-09-05`) %>% 
    select(-contains("not"))
    

```

```{r}
tm_shape(NYC_FEMA_Spatial) + 
    tm_polygons(col='Eligible_2012-10-30', title = "Sandy FEMA Eligible Assistances", palette = "viridis") + 
    tm_scale_bar(position = c("right", "top")) 

tm_shape(NYC_FEMA_Spatial) + 
    tm_polygons(col='Ineligible_2012-10-30', title = "Sandy FEMA Ineligible Assistances", palette = "viridis") + 
    tm_scale_bar(position = c("right", "top")) 

tm_shape(NYC_FEMA_Spatial) + 
    tm_polygons(col='Eligible_2021-09-05', title = "Ida FEMA Eligible Assistances", palette = "viridis") + 
    tm_scale_bar(position = c("right", "top")) 

tm_shape(NYC_FEMA_Spatial) + 
    tm_polygons(col='Ineligible_2021-09-05', title = "Ida FEMA Ineligible Assistances", palette = "viridis") + 
    tm_scale_bar(position = c("right", "top")) 
```
```{r}

tm_shape(NYC_FEMA_Spatial) + 
    tm_polygons(col='eligible_sandy_pc', title = "Sandy FEMA Eligible Assistances", palette = "viridis", style = "jenks") + 
    tm_scale_bar(position = c("right", "top")) 

tm_shape(NYC_FEMA_Spatial) + 
    tm_polygons(col='ineligible_sandy_pc', title = "Sandy FEMA Ineligible Assistances", palette = "viridis",style = "jenks") + 
    tm_scale_bar(position = c("right", "top")) 

tm_shape(NYC_FEMA_Spatial) + 
    tm_polygons(col='eligible_Ida_pc', title = "Ida FEMA Eligible Assistances", palette = "viridis",style = "jenks") + 
    tm_scale_bar(position = c("right", "top")) 

tm_shape(NYC_FEMA_Spatial) + 
    tm_polygons(col='ineligible_Ida_pc', title = "Ida FEMA Ineligible Assistances", palette = "viridis",style = "jenks") + 
    tm_scale_bar(position = c("right", "top")) 
```
```{r}
tm_shape(NYC_FEMA_damage_Spatial) + 
    tm_polygons(col='damaged_sandy', title = "Sandy FEMA damage", palette = "viridis",style = "jenks") + 
    tm_scale_bar(position = c("right", "top")) 

tm_shape(NYC_FEMA_damage_Spatial) + 
    tm_polygons(col='damaged_sandy_pc', title = "Sandy FEMA damages per capita", palette = "viridis",style = "jenks") + 
    tm_scale_bar(position = c("right", "top")) 

tm_shape(NYC_FEMA_damage_Spatial) + 
    tm_polygons(col='damaged_ida', title = "Ida FEMA damage", palette = "viridis",style = "jenks") + 
    tm_scale_bar(position = c("right", "top")) 

tm_shape(NYC_FEMA_damage_Spatial) + 
    tm_polygons(col='damaged_ida_pc', title = "Ida FEMA damages per capita", palette = "viridis",style = "jenks") + 
    tm_scale_bar(position = c("right", "top")) 
```


```{r}
write_sf(NYC_FEMA_Spatial, "../3_Intermediate/FEMA_Cleaned.shp")
write_sf(NYC_FEMA_damage_Spatial, "../3_Intermediate/FEMA_Cleaned_damage.shp")
## Write to SF, then bring back to FEMA_Exploration
```



