---
title: "Data_Exploration"
author: "Sean"
date: "1/17/2022"
output:
  html_document:
    df_print: paged
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
### Set Environment ------------------------------------------------------------
# Clear the environment
#rm(list=ls())
gc()
# So the code will compile warnings as per usual
options(warn = 0)
# Turn off scientific notation
options(scipen = 999)
### Load Packages --------------------------------------------------------------
if(!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, magrittr, stringr, reshape2, janitor,
lubridate, readxl, data.table, ggplot2, scales, readr,
tidyr, zoo, skimr, openxlsx,ggspatial, rgeos, data.table,RColorBrewer,
tidyverse, rio, collapse, sf, glue, XML, tm, here, purrr, repurrrsive,
tmap,tidygraph, nabor,igraph, viridis, hrbrthemes,RSocrata,soql,xts,
tidycensus)
```


### Functions Used:
```{r}
#source("./Exploration_Functions.R")

```

# Data Exploration of 311 Data

## Part 1: Exploring different agencies and setting API up:

### First Test of 311 data:

```{r}
dataset_id <- "erm2-nwe9" #311 complaint open data
api_endpoint <- paste0('https://data.cityofnewyork.us/resource/', dataset_id, '.json')
api_token <- "YRWk8RdZZ9xPMVbxc0QaHgYFU"

parks_id <- "y6ja-fw4f"
parks_endpoint <- paste0('https://data.cityofnewyork.us/resource/', parks_id, '.json')
```
First step will be to set up our environment and begin querying using the 
socrata package. I will test 10,000 entries first.

```{r}
query_params <- soql() %>%
    soql_add_endpoint(api_endpoint) %>%
    soql_limit(10000)

query <- read.socrata(
  query_params,
  app_token = api_token
)

names(query)
agencies <- unique(query$agency)
agencies
```


### Exploring different agencies:
Because there are many agencies, I will need to dive in to see what agencies 
there are. Agencies I'm interested in include Department of Environmental 
Protection, Emergency Management, Health and Mental Hygenie, 311, and buildings
```{r}

descriptors <- function(department) {
    descriptors <- query %>% 
        filter(agency == department) %>% 
        pull(descriptor) %>% 
        unique() %>% 
        as.data.frame()
    return(descriptors)
}

agencies <- c("DOHMH","DEP","DOB")

DOHMH <- descriptors("DOHMH")
DEP <- descriptors("DEP")
DOB <- descriptors("DOB")


DOHMH
DEP
DOB


```


## Part 2: Diving into DEP data:

### Pulling DEP 311 complaints:
Because there are many entries, I created an API token to not encounter throttle
limits. There are millions of entries, so I will save in a more compact data 
type, RDS. We can load this in quicker next time. The Query will be commented
out so we can query again if needed.
```{r}
## For Querying:
DEP_params <- soql() %>%
    soql_add_endpoint(api_endpoint) %>%
    soql_simple_filter("agency","DEP") %>% 
    soql_limit(100)

DEP_Query <- read.socrata(
  DEP_params,
  app_token = api_token
)
        
# saveRDS(DEP_Query,"DEP_Data.rds")

DEP_Query <- read_rds("../3_Intermediate/DEP_Data.rds")
skim(DEP_Query)
```

### Looking at the most important complaint types:
Overall, it looks like the highest complaints are from water system, followed by
Noise, and Sewer. 
```{r}
DEP_Query_by_type <-DEP_Query %>% 
  group_by(complaint_type) %>% 
  count() %>% 
  arrange(by = n)

DEP_Query_by_type
```

```{r}
head(DEP_Query)
IDA_Data <-  DEP_Query %>% 
    filter(created_date >= as.Date("2021-08-26")) %>% 
    filter(created_date <= as.Date("2021-09-04")) 

IDA_Date <- IDA_Data%>% 
    mutate(day = day(created_date),
           month = month(created_date),
           year = year(created_date)) %>% 
    group_by(complaint_type,year,month,day) %>% 
    count() %>% 
    mutate(date = as.Date(paste0(year,"-",month,"-",day)))

names(IDA_Date)

ggplot(IDA_Date, aes(IDA_Date$date, n ,color = IDA_Date$complaint_type)) +
    geom_line() +
    scale_color_viridis(discrete = T) +
    theme_minimal() +
    xlab("Ida Timeframe") +
    scale_y_continuous(label=comma)

#ggsave("../4_Outputs/IDA_complaints.png")

Sandy_Data <-  DEP_Query %>% 
    filter(created_date >= as.Date("2012-10-22")) %>% 
    filter(created_date <= as.Date("2012-11-2")) 

Sandy_Date <- Sandy_Data %>% 
    mutate(day = day(created_date),
           month = month(created_date),
           year = year(created_date)) %>% 
    group_by(complaint_type,year,month,day) %>% 
    count() %>% 
    mutate(date = as.Date(paste0(year,"-",month,"-",day)))

ggplot(Sandy_Date, aes(Sandy_Date$date, n ,color = Sandy_Date$complaint_type)) +
    geom_line() +
    scale_color_viridis(discrete = T) +
    theme_minimal() +
    xlab("Sandy Timeframe") +
    scale_y_continuous(label=comma)

#ggsave("../4_Outputs/Sandy_complaints.png")

```


```{r}
DEP_points <- read_rds("../3_Intermediate/DEP_points.rds")

IDA_Data <-  DEP_points %>% 
    filter(created_date >= as.Date("2021-08-26")) %>% 
    filter(created_date <= as.Date("2021-09-04"),
           complaint_type == "Sewer") %>%
    mutate(day = day(created_date),
           month = month(created_date),
           year = year(created_date),
           date = as.Date(paste0(year,"-",month,"-",day)))

Sandy_Data <- DEP_points %>% 
    filter(created_date >= as.Date("2012-10-22")) %>% 
    filter(created_date <= as.Date("2012-11-2"),
           complaint_type == "Sewer") %>% 
    mutate(day = day(created_date),
           month = month(created_date),
           year = year(created_date),
           date = as.Date(paste0(year,"-",month,"-",day)))

```

```{r}
# boroughs <- read_sf("../2_Data/borough_boundaries/geo_export_b5bd9a81-0a45-411a-a23f-8a1983296a03.shp")
tracts <- read_sf("../2_Data/ct_2010/geo_export_1c19ce5f-d77c-4a3f-adbe-7456e4a782a6.shp")
parks <- read_sf("../2_Data/Open Space (Parks)/geo_export_2addf878-38fa-4c00-8396-5aa5bf54644c.shp")


# tracts <- get_acs(geography = "tract",
#               state = "New York",
#               county = c(
#                 "Bronx",
#                 "Kings",
#                 "New York",
#                 "Queens", 
#                 "Richmond"),
#               year = 2020,
#               variables = "B19013_001",
#               key = "00c9ebc1f593573f959d804c31974d714e280b35",
#               geometry = TRUE)
# 
# park_params <- soql() %>%
#     soql_add_endpoint(parks_endpoint) 
# 
# parks_Query <- read.socrata(
#   park_params,
#   app_token = api_token
# )

```

```{r}
anim_ida <- tm_shape(tracts) +
    tm_polygons(lwd = .1) +
    tm_shape(parks) +
    tm_fill(col = "#99d8c9",alpha = .5)+
    tm_shape(IDA_Data)+
    tm_dots() +
    tm_facets(along = "date", free.coords = FALSE) +
    tm_layout(
              main.title.size = 1,
              main.title.fontface = 2
    )

anim_ida
```


```{r}
tmap_animation(tm=anim_ida, filename = "../4_Outputs/Ida.gif", delay = 100)
```


```{r}

anim_sandy <- tm_shape(tracts) +
    tm_polygons(lwd = .1) +
    tm_shape(parks) +
    tm_fill(col = "#99d8c9",alpha = .5)+
    tm_shape(Sandy_Data)+
    tm_dots() +
    tm_facets(along = "date", free.coords = FALSE) +
    tm_layout(
              main.title.size = 1,
              main.title.fontface = 2
    )

anim_sandy


```


```{r}
tmap_animation(tm=anim_sandy, filename = "../4_Outputs/sandy.gif", delay = 100)
```






